From Trends to Truth: A Comprehensive Analysis Revealing Contributing Factors Behind Declining US Birth Rates

### 2024-02-14 paper/paper.qmd
fig 4 result section

@fig-1 showcases the trajectory of US birth rates spanning from 1980 to 2020, focusing on births per 1000 women aged 15 to 44. The peak occurred in 1990, reaching 70.9 births per 1000 women, closely followed by 2007 with 69.3 births per 1000 women. In contrast, 2020 marked the lowest point with only 55.8 births per 1000 women. From 1980 to 2007, the birth rate experienced moderate fluctuations, with a maximum variance of 7.3 births per 1000 women. However, following the year 2007, the birth rate seems to display a consistent a downward trend, declining by 13.5 births per 1000 women between 2007 and 2020.

@@

Examining the trends in birth rates by five-year age groups reveals intriguing patterns.

@@

Unlike the figures discussed previously, @fig-2b examines US birth rate trends concerning the mother's level of education, focusing on the period from 1992 to 2019 and women aged 20 to 44.

@@

>>>>

@fig-1 is the replication of Figure 1 from `The Puzzle of Falling US Birth Rates since the Great Recession' by @citeOriginalReplication, and showcases the trajectory of US birth rates spanning from 1980 to 2020, focusing on births per 1000 women aged 15 to 44. The peak occurred in 1990, reaching 70.9 births per 1000 women, closely followed by 2007 with 69.3 births per 1000 women. In contrast, 2020 marked the lowest point with only 55.8 births per 1000 women. From 1980 to 2007, the birth rate experienced moderate fluctuations, with a maximum variance of 7.3 births per 1000 women. However, following the year 2007, the birth rate seems to display a consistent a downward trend, declining by 13.5 births per 1000 women between 2007 and 2020.

@@

@fig-2a is the replication of Figure 2 (a) from the original paper. Examining the trends in birth rates by five-year age groups reveals intriguing patterns.

@@

@fig-2b is the replication of Figure 2 (d) from the original paper, and examines US birth rate trends concerning the mother's level of education, focusing on the period from 1992 to 2019 and women aged 20 to 44.

@@

@fig-2c replicates figure 2 (f) from @citeOriginalReplication, and breaks down the trends in birth rates by parity.

In this case, all subgroups representing first, second, third, and fourth or more births exhibit nearly identical patterns, maintaining a stable birth rate. However, there is a notable distinction: the fourth or more birth subgroup shows slight growth from around 1995 to 2019, while all other parity groups witness a decline in births per 1000 women from approximately 2007 to 2019.

Consistently, the first birth subgroup boasts the highest number of births per 1000 women, followed by the second birth subgroup. Conversely, the fourth or more birth subgroup, despite displaying a positive trend from 1995 to 2000, consistently records the lowest number of births per 1000 women each year, followed by the third birth subgroup.
